Argennon: A Scalable Cloud Based Smart Contract Platform Using Argument of Knowledge Systems

### 2022-09-20 tex/sections/consensus.tex
Complete the new emergency forking protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The merging will continue recursively until the online stake of all remaining assemblies is higher than
\texttt{minOnlineStake}. If eventually all assemblies get merged together and only one assembly remains, the condition
for validity of block certificates changes: A certificate of validators will be considered valid if the sum of stakes
of the certificate signers is higher than $2/3$ of the total stake of validators and \texttt{online/offline}
status of validators becomes ineffective.

@@

In particular we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
if it is signed by more than $\frac{2}{3}$ of the total stake of the assembly and there is no \texttt{online/offline}
status for users.

@@

The recovery protocol can recover the functionality
of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
agreement} protocol.

@@

The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
from the validators.
This fork needs to be confirmed by validators and can not discard any blocks that has been already certified
by validators.
A valid fork request is an unexpired request signed by more than half of the total stake of validators.

For forking at block $h$, the reserve committee of delegates
makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $h$.
The height of the fork block therefore is
$h + 1$, so it needs a valid certificate from the assembly of validators with the index $h+1$ modulo $m$.

For signing a fork block, a validator ensures that the block is signed by the reserve committee, contains
a valid fork request and its parent is not issued by the same reserve committee. The parent of the fork block does not
necessarily need a validators certificate. If the parent does not have a certificate, the validator checks the
certificate of the block before the parent and the conditional validity of the parent instead. This enables the
reserve committee to recover the liveliness of the blockchain in a situation where a malicious committee has
generated multiple blocks at the same height. Notice that the block before the parent always need a validators
certificate.

A validator always chooses a valid fork block over a block of the main chain. However, as we mentioned before,
a honest validator \textbf{never signs a certificate for two different blocks with the same height}. Consequently, if a
validator has already signed the block $h+1$ of the main chain, he will not sign the fork block at height $h+1$ and
vice versa.

The reserve committee of delegates is not allowed to generate multiple fork blocks at the same height. However, they
can generate multiple fork blocks with different heights, as long as the parent block is generated by the previous
committee. When the reserve committee generates a fork block at height $h+1$, but the block fails to receive a
certificate from validators, the committee will generate a new fork block at height $h+2$.

Moreover, an honest committee of delegates will always try to perform the emergency forking in such a way that
valid blocks do not get discarded, including blocks that are not certified by the validators yet.

@@

When the reserve committee gets activated, the main committee might have been malicious, so any number of blocks could
exist at each height. However, at each height at max one block can have a validators certificate. Moreover, if at some
height there are not any blocks with a validators certificate, then no blocks at higher heights can have a
validators certificate, because validators do not sign the certificate of a normal block if its parent does
not have a certificate.

If $h_{\max}$ denotes the height of the highest block with a validators certificate and a fork block is generated at
height $h$, as long as more than $2/3$ stake of all the assemblies of validators are honest, we have:
\begin{itemize}
    \item if $h > h_{\max} + 2$, the fork block can not receive a certificate from validators.
    \item if $h = h_{\max} + 2$, the fork block can always receive a certificate from validators if network
    partitions do not last forever.
    \item if $h = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
    the validators of the assembly with index $h+1$ modulo $m$ get divided between the fork block and a block at
    height $h+1$ of the main chain.
    \item if $h < h_{\max} + 1$, the fork block can not receive a certificate from validators.
\end{itemize}

Obviously, if the reserve committee generates two fork blocks at height $h_{\max} + 2$ and $h_{\max} + 1$, both blocks
could receive a validators certificate. However, this will not be problematic, since when the fork block at height
$h_{\max} + 2$ has a validators certificate, any block added after the fork block at height $h_{\max} + 1$ can not
receive the certificate of validators and the fork block at this height is effectively sealed.

@@

\newpage
A user executes the following procedure in each agreement session:
\paragraph{Voting Phase in Round $r$:}
\begin{itemize}
    \item if the user has final-voted any value, he $r$-votes his final-vote, otherwise he $r$-votes a single desired
    proposal.
    \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
    \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
    \item if $clock > r \cdot \lambda$ and $r < k$ user goes to the round $r + 1$ and if $r = k$ user starts a
    new agreement session.
\end{itemize}

>>>>

The merging will continue recursively until the online stake of all remaining assemblies is higher than
\texttt{minOnlineStake} fraction.
If eventually all assemblies get merged together and only one assembly remains, the condition
for validity of block certificates changes: A certificate of validators will be considered valid if the sum of stakes
of the certificate signers is higher than $2/3$ of the total stake of validators and \texttt{online/offline}
status of validators becomes ineffective.

@@

In particular, we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
if it is signed by more than $\frac{2}{3}$ of the total stake of the assembly and there is no \texttt{online/offline}
status for users.

@@

We assume that an adversary is able to fully control message transmission between users and is able to partition the
network arbitrarily for finite periods of time. Under these circumstances, the recovery protocol can recover the
functionality of the Argennon blockchain as long as more than $2/3$ of the total stake of every validator assembly is
controlled by honest users. The recovery protocol uses two main emergency procedures to
recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
agreement} protocol.

@@

The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
from the validators.
This fork needs to be confirmed by validators and can not discard more than one block that has been certified by
validators.
A valid fork request is an unexpired request signed by more than half of the total stake of validators.

For forking at block $h$, the reserve committee of delegates
makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $h$.
The height of the fork block therefore is $h + 1$, and the fork block needs a valid certificate from the assembly of
validators with the index $h+1$ modulo $m$. When a
fork block gets certified by validators, its parent is also confirmed and will become a part of the blockchain, even if
it does not have a validators certificate.

For signing a fork block at height $h+1$, a validator ensures that the following conditions hold:
\begin{itemize}
    \item the fork block is signed by the reserve committee.
    \item the fork block contains a valid fork request.
    \item the parent block of the fork block is issued by the previous committee.
    \item the parent block of the fork block is certified by validators, or the parent block is conditionally
    valid and there is a fork block with height $h$ which is certified by validators, or the parent block is
    conditionally valid and the parent block of the parent has a validators certificate.
    \item the validator has not already signed a certificate for a fork or normal block at height $h+1$.
\end{itemize}

The parent of the fork block does not necessarily need a validators certificate. This enables the
reserve committee to recover the liveliness of the blockchain in a situation where a malicious committee has
generated multiple blocks at the same height. Notice that the block before the parent always needs a validators
certificate.

A validator always chooses a valid fork block over a block of the main chain and may sign different fork
blocks with different heights. However, as we mentioned before, an honest validator
\textbf{never signs a certificate for two different blocks with the same height}. Consequently, a validator never
signs two fork blocks at the same height, and if he has already signed the fork block at height $h+1$, he will not
sign the block $h+1$ of the main chain and vice versa.

The reserve committee of delegates is allowed to generate multiple fork blocks with different heights, as long as the
parent block is generated by the previous committee. When the reserve committee generates multiple fork blocks
at different heights, the next normal block must be always added after the fork block with the highest height.

The reserve committee of delegates should try to perform the emergency forking in such a way that
valid blocks do not get discarded, including blocks that have not been certified by the validators yet.

For forking the blockchain, the reserve committee uses a straightforward algorithm. Let $h_v$ be the height of the last
block with a validator certificate and $h_v+k$ be the height of the last valid block that the reserve committee has
seen. For forking the main chain, the reserve committee generates all fork blocks with heights $h_v+1,h_v+2,\dots,
h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block $h_v+i-1$ of the main chain. Then the
reserve committee will wait until the fork block with height $h_v+k+1$ receives a certificate form validators. The
reserve committee then will continue the normal chain after that fork block and the fork block with height $h_v+k+1$
will be the parent of the first normal block generated by the reserve committee.

@@

When the reserve committee gets activated, the main committee might have been malicious, so any number of blocks could
exist at each height. However, at each height at max one block can have a validators certificate. Moreover, if at some
height there are not any blocks with a validators certificate, then no blocks at higher heights can have a
validators certificate either, because validators do not sign the certificate of a normal block if its parent does
not have a certificate.

If $h_{\max}$ denotes the height of the highest block with a validators certificate, as long as more than $2/3$ stake
of all the assemblies of validators are honest, for a fork block with height $h_f$ we have:
\begin{itemize}
    \item if $h_f \leq h_{\max}$, the fork block can not receive a certificate from validators.
    \item if $h_f = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
    the validators of the assembly with index $h_f$ modulo $m$ get divided between the fork block and a block at
    height $h_f$ of the main chain.
    \item if $h_f = h_{\max} + 2$, the fork block can always receive a certificate from validators, if network
    partitions do not last forever.
    \item if $h_f \geq h_{\max} + 3$, the fork block can always receive a certificate from validators \textbf{only if} a
    fork block at height $h_f-1$ gets certified by validators.
\end{itemize}

Consequently, if two fork blocks at heights $h_0$ and $h_0+k$ are generated by the reserve committee, if both blocks
receive a certified from validators, we must have $h_0 > h_{\max}$, and there must exist fork blocks with heights
$h_0+1,\dots,h_0+k-1$ which are certified by validators. That means if a malicious reserve committee generates a normal
block after any of the fork blocks with heights less than $h_0+k$, that normal block can not receive a certificate from
validators.

@@

A user executes the following procedure in each agreement session:
\paragraph{Voting Phase in Round $r$:}
\begin{itemize}
    \item if the user has final-voted any value, he $r$-votes his final-vote, otherwise he $r$-votes a single desired
    proposal.
    \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
    \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
    \item if $clock > r \cdot \lambda$ and $r < k$ user goes to the round $r + 1$ and if $r = k$ user starts a
    new agreement session.
\end{itemize}
